Behavior Driven Development in Haskell

### 2012-06-02 paper.tex
More conservative wording

We present a step-by-step refinement of a library
that, and a corresponding spec.  Each line of code is
only given once when introduced, subsequent code fragments are
incremental updates.  The reader is asked to substitute code parts as
appropriate.

\subsubsection{Start with a minimal implementation}

@@

\subsubsection{Get it right}

In the next step, we add a specification that forces our function to
actually calculate Fibonacci numbers.

\begin{quote}
    TBD: do we want to add something like this?

    At this stage we do not think about invalid input or performance.
    We are happy with an implementation that works for small
    non-negative integers.
\end{quote}

\begin{quote}
\small
\begin{verbatim}
-- small non-negative numbers
newtype Small = Small Int
  deriving Show

instance Arbitrary Small where
  arbitrary = Small . (`mod` 10) <$> arbitrary
\end{verbatim}
\end{quote}

\noindent TBD: can we put some useful text here, so that the code
example is divided into two pieces, and hence each piece is smaller?

\begin{quote}
\small
\begin{verbatim}
main = hspec spec

@@

\noindent Of course this spec fails when we run it,

@@

\noindent so we add a working implementation.  Both Doctest and Hspec
report success after that.

@@

\subsubsection{Get it fast}

As the seasoned Haskeller might have already noticed, our
implementation is quite inefficient, as made evident by this
specification:

@@

\noindent When we run it, hspec tells us

@@

\noindent and we respond with an efficient implementation.

@@

\noindent TBD: With a much more efficient implementation at hand, we
can also adapt our definition of ``small non-negative number'', so
that a wide range of input values is tested:

@@

So far, so good.  We have an implementation of the Fibonacci function
that swiftly calculates correct values for all non-negative integers.
But what about negative ones? Let's assume that we want our
implementation to throw an exception in this case. We would
specify it like this.

@@

Some time later however, the one person in our team that knows
more about Haskell than we do, convinces us that throwing exceptions
from pure code is considered poor design. Hence, we change the
specification so that negative inputs should have zero as result.

>>>>

We present a step-by-step refinement of a library, and a corresponding
spec.  Each line of code is only given once when introduced,
subsequent code fragments are incremental updates.  The reader is
asked to substitute code parts as appropriate.

\subsubsection{Starting with a minimal implementation}

@@

\subsubsection{Getting it right}

Next, we add a specification that forces our function to actually
calculate Fibonacci numbers.
At this stage we do not worry about invalid input or performance.  We
are happy with an implementation that works for small non-negative
integers.

\begin{quote}
\small
\begin{verbatim}
newtype Small = Small Int
  deriving Show

instance Arbitrary Small where
  arbitrary = Small . (`mod` 10) <$> arbitrary

@@

\noindent This spec fails when we run it:

@@

\noindent So we add a working implementation:

@@

\noindent Both Doctest and Hspec report success now.

\subsubsection{Getting it fast}

Our implementation is quite inefficient, as made evident by this
specification:

@@

\noindent When we run it, hspec tells us:

@@

\noindent And we respond with an efficient implementation:

@@

\noindent With a much more efficient implementation at hand, we can
now adapt our definition of small non-negative integers, so that
a wider range of input values is tested:

@@

We have an implementation of the Fibonacci function
that calculates correct values for all non-negative integers.
But what about negative ones? Let's assume that we want our
implementation to throw an exception in this case. We would
specify it like this.

@@

Some time later however, the one person in our team that knows
more about Haskell than we do, convinces us that throwing exceptions
from pure code is considered poor design. Hence, we change the
specification so that negative input values should result in zero.
